Uncertainty in social learning evolution

### 2022-05-22 ms.tex
updated figures; makefile

To answer our research questions
we developed an agent-based model of social learning evolution in which a society of
$N$ individuals each must decide which of $B$ behaviors to perform at each time step
within a generation of $L$ time steps. After each generation agents are selected
with replacement from the population to reproduce another $N$ agents. Agents who
inherit the social learning trait then learn about behavioral payoffs from the
previous generation, while asocial learners begin life with no \emph{a priori}
knowledge of the world.  Each behavior is a ``bandit'', a common modelling and
experimental approach for representing behaviors with probabilistic
payoffs~\cite{SuttonBartoBook,McElreath2005,Rendell2010,Schulz2019}.  Each behavior
$b$ yields Bernoulli payoffs: payoff of 1 with probability $\pi_b$ and zero payoff
with probability $1 - \pi_b$; $\pi_b$ is therefore also the expected payoff of
behavior $b$. One behavior in each generation is optimal and pays off $\pihigh$,
while the rest pay off $\pilow < \pihigh$.  Model agents must decide which behavior to
perform at each time step.  To do this, agents use an explore-exploit strategy to
sometimes try the most profitable behavior they know about, and other times try
alternatives that may pay off more reliably.  We developed a series of computational
analyses by systematically varying uncertainty parameters and observing the
frequency that model populations evolve to be social learners, the average payoffs
accrued by agents in each setting, and the time it takes for model populations to
reach fixation (agents all social or all asocial learners).

@@

\paragraph{Analysis 1: patterns in social learning evolution over $u$}

@@

>>>>

To answer our research questions
we developed an agent-based model of social learning evolution in which a society of
$N$ individuals each must decide which of $B$ behaviors to perform at each time step
within a generation of $L$ time steps. At the end of each generation, agents are selected
with replacement from the population to reproduce another $N$ agents. Agents who
inherit the social learning trait then learn about behavioral payoffs from the
previous generation, while asocial learners begin life with no \emph{a priori}
knowledge of the world.  Each behavior is a ``bandit'', a common modeling and
experimental approach for representing behaviors with probabilistic
payoffs~\cite{SuttonBartoBook,McElreath2005,Rendell2010,Schulz2019}.  Each behavior
$b$ yields Bernoulli payoffs: payoff of 1 with probability $\pi_b$ and zero payoff
with probability $1 - \pi_b$; $\pi_b$ is therefore also the expected payoff of
behavior $b$. One behavior in each generation is optimal and pays off $\pihigh$,
while the rest pay off $\pilow < \pihigh$.  Model agents must decide which behavior to
perform at each time step.  To do this, agents use an explore-exploit strategy to
sometimes try the most profitable behavior they know about, and other times try
alternatives that may pay off more reliably.  We developed a series of computational
analyses by systematically varying uncertainty parameters and observing the
frequency that model populations evolve to be social learners, the average payoffs
accrued by agents in each setting, and the time it takes for model populations to
reach fixation (agents all social or all asocial learners).

@@

\begin{figure}
  \caption{Social learning prevalence (y-axes) monotonically decreases as 
  environmental variabilty, $u$, increases (x-axes) in most uncertainty contexts. 
  Other uncertainty values $\pilow$ (rows), $B$ (columns), and $L$ (keys)
  shift and flatten the decrease from all-social-learner populations to all-asocial 
  populations.}
  \label{fig:mainResults}
  \centering
    \includegraphics[width=\textwidth]{Figures/mainResultsPlots.pdf}
\end{figure}

@@

\begin{figure}
  \caption{Mean payoff normalized for lifespan (y-axes, solid line with circles)
    generally track either expected all-social-learners payoff ($\meansoc$, dash-dot
    line with diamonds) or expected all-asocial payoff ($\meanasoc$, long dash
    horizontal lines) as $u$ increases (x-axes). When $\meansl$ goes from 1 to
  0 is when social learning payoffs approach individual learning payoffs due to
increased environmental variability.} 
  \label{fig:netPayoffs}
\centering
    \includegraphics[width=0.75\textwidth]{Figures/meanNetPayoffs.pdf}
\end{figure}

\begin{figure}
  \caption{Average number of generations ($\meanT/L$, y-axes) for populations to fixate
  where agents either all become social or all become asocial learners. As
  $u$ increases (x-axes) models take longer to fixate due to drift, up to a point,
but then fixation accelerates as individual learning becomes more obviously 
superior. Increased $\pilow$ increases time to fixate; both $\pilow$ and $B$
shift peak time to fixation over $u$.} 
  \label{fig:steps}
\centering
    \includegraphics[width=0.75\textwidth]{Figures/stepResultsPlots.pdf}
\end{figure}
